Biologically-Constrained 3D Agglomeration for Connectome Reconstruction

### 2018-03-14 papers/eccv2018/1_related_work.tex

% \begin{figure}[t]
% 	\subfloat[]{
% 		\includegraphics[width=.47\linewidth]{figures/schema/pre-multicut-with-arrows.pdf}
% 	}
% 	\subfloat[]{
% 		\includegraphics[width=.47\linewidth]{figures/schema/post-multicut.png}
% 	}
% 	\caption{Example improvement of neural reconstruction. (a) We extract 3D skeletons from pixel-based segmentation algorithms to create a graph representation. Red arrows highlight errors in the input segmentation. (b) We improve the segmentation accuracy using a graph partitioning algorithm, leveraging both local and global information.}
% 	\label{fig:improved-reconstruction}
% \end{figure}

\subsubsection{Initial Segmentation}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level from raw EM images.

First, intermediate representations like boundary or affinity maps are generated. Some methods use a random forest populated with hand-designed features based on image characteristics around a given pixel to predict membrane probabilities~\cite{kaynig2015large}.
Most recent advancements use convolutional neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
Originally, networks predicted membrane probabilities per each image slice using 2D convolutions~\cite{seymour2016rhoananet,ronneberger2015u,ciresan2012deep,jain2010boundary,kaynig2015large,amelio_segmentation}.
However, recent research extends these convolutions to include the z-dimension improving performance~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional}.
Many of these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). 
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. 
These networks produce impressive accuracies but at a high computational cost.
Data augmentation in both training and testing can increase the prediction accuracy in boundary detection~\cite{zeng2017deepem3d}.
Lee et al. surpass the estimated human accuracy on the SNEMI3D dataset with a variant of the 3D U-Net architecture~\cite{lee2017superhuman}.

Then, these intermediate representations are transformed into segmentation.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}.
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.

\subsubsection{Agglomeration}
Given the initial segmentation, some methods train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.
\TODO{add learning-based algorithm} ~\cite{jain2011learning,bogovic2013learned}
%\subsubsection{Error-correction methods}
%Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
%However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
%This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

>>>>

\begin{figure}[t]
 	\subfloat[]{
 		\includegraphics[width=.47\linewidth]{figures/schema/pre-multicut-with-arrows.pdf}
 	}
 	\subfloat[]{
 		\includegraphics[width=.47\linewidth]{figures/schema/post-multicut.png}
 	}
 	\caption{Example improvement of neural reconstruction. (a) We extract 3D skeletons from pixel-based segmentation algorithms to create a graph representation. Red arrows highlight errors in the input segmentation. (b) We improve the segmentation accuracy using a graph partitioning algorithm, leveraging both local and global information.}
 	\label{fig:improved-reconstruction}
\end{figure}

\subsubsection{Pixel-Based Segmentation}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level from raw EM images. First, intermediate representations like boundary or affinity maps are generated. Some methods use a random forest populated with hand-designed features based on image characteristics around a given pixel to predict membrane probabilities~\cite{kaynig2015large}.
Most recent advancements use convolutional neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
Originally, networks predicted membrane probabilities per each image slice using 2D convolutions~\cite{seymour2016rhoananet,ronneberger2015u,kaynig2015large,ciresan2012deep,jain2010boundary,amelio_segmentation}.
More recent research extends these convolutions to include the z-dimension improving performance~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional}.
Many of these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). 
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. 
These networks produce impressive accuracies but at a high computational cost.
Data augmentation in both training and testing can increase the prediction accuracy in boundary detection~\cite{zeng2017deepem3d}.
Lee et al. surpass the estimated human accuracy on the SNEMI3D dataset with a variant of the 3D U-Net architecture~\cite{lee2017superhuman}.

\subsubsection{Region Agglomeration}
Next, agglomeration techniques transform these intermediate representations into segmentations.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}.
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster these pixels into super-pixels~\cite{zlateski2015image}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Given the initial segmentation, others train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}
extract different features~\cite{bogovic2013learned,jain2011learning}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.
In our work, we combine these last two strategies of classification and optimization.

\subsubsection{Error-correction methods}
Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
We extract a 3D graph from an input segmentation for a true top-down error correction approach. 
This allows us to enforce domain-specific biological constraints while using graph partitioning algorithms. 
We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

%\subsubsection{Error-correction methods}
%Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
%However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
%This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

### 2018-03-14 papers/eccv2018/4_results.tex

\subsection{Connectomics Benchmark}
In Fig.~\ref{fig:variation-of-information}, we show the VI results of the pixel-based reconstructions of the Kasthuri and FlyEM data (Sec.~\ref{sec:neuroproof}) for varying thresholds of agglomeration (green). 
We use one of these segmentations (green circle) as our input dataset with an agglomeration threshold of 0.3 for all datasets. 
The results from our method are shown in red for varying the $\beta$ parameter. 
We show comparisons to an oracle (blue) that correctly partitions the graph from our method based on ground truth.

Our algorithm improves the accuracy of the reconstruction for every dataset, reducing the VI split score on average by \TODO{XX}\% on the three testing datasets. 
Scores closer to the origin are better for this metric, and in every instance our results are below the green curve.
We see significant improvements on the Kasthuri datasets (VI split reduction of \TODO{XX}\%) and more modest improvements on the FlyEM dataset (reduction of \TODO{XX}\%) and Fib-25 dataset (reduction of \TODO{XX}\%).

@@

\subsection{Empirical Ablation Studies}
Below, we analyze various design choiceswe for each step of our pipeline.

\noindent\textbf{Graph Generation.}
Table \ref{table:skeletonization} shows the results of pruning the skeleton graph using the algorithm discussed in Sec.~\ref{sec:skeletonization}. 
This edge pruning is essential for the graph partitioning algorithm, which has a computational complexity dependence on the number of edges. 
The baseline algorithm considers all adjacent regions for merging. 
Our method removes a significant portion of these candidates while maintaining a large number of the true merge locations (e.g., \TODO{XXX} compared to \TODO{XXX}). 
Our pruning heuristic removes at least $\TODO{X}\times$ the number of edges on all datasets, achieving a maximum removal rate of $\TODO{XX}\times$.
However, there are some adjacent over-segmented labels which are not considered. 
Table~\ref{table:skeletonization} shows the results for our edge generation including the number of missed neighbors and potential penalty in VI score.
We present additional analysis of the skeletonization parameters in the supplemental material.

\begin{table}
	\centering
	\small
	\setlength{\tabcolsep}{0.75em}
	\begin{tabular}{c c c c c} \hline
		\textbf{Dataset} & \textbf{Segment Adjacency} & \textbf{Skeleton Pruning} & \textbf{Missed Splits} & \textbf{Gained Edges} \\ \hline
		Kasthuri & 610 / 23506 & 556 / 4921 & 214 & 160 \\
		FlyEM & 298 / 17614 & 197 / 4366 & 120 & 19 \\
		FIB-25 &  & & & \\ \hline
	\end{tabular}
	\caption{The results of our graph pruning approach compared to the baseline graph with all adjacent regions. We show the number of true merge locations (e.g., 610) compared to total number of edges in the graph (e.g., 23506) for each case. The number of missed splits corresponds to the number of split errors that our method misses compared to an adjacency matrix.}
	\label{table:skeletonization}
\end{table}

We generate edges in our graph by using information from the skeletons. 
In particular, we do not enforce the constraint that edges in our graph correspond to adjacent segments.
Although neurons are continuous, the EM images often have noisy spots which cause an interruption in the input segmentation.
We still want to reconstruct these neurons despite the fact that the initial segmentation is non-continuous. 
The second and fourth examples in Fig.~\ref{fig:qualitative-results} show correctly reconstructed neurons where two of the segments are non-adjacent. 
On average in our datasets we correctly merge \TODO{XX} number of non-adjacent segments.

There are some pairs of segments which we do not consider for merging because of our reliance on the skeletons.
Fig.~\ref{fig:skeleton-results} shows two such cases. 
The endpoints of both segments are circled.
In this example the small segment is carved from the larger segment in a location where there are no skeleton endpoints. 
There are on average $\TODO{XX}$ such examples in our datasets, corresponding to $\TODO{XX}$ in the variation of information split score.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\linewidth]{./figures/receiver-operating-characteristic.png}		\includegraphics[width=0.45\linewidth]{./figures/precision-and-recall.png}
	\caption{The receiver operating characteristic (ROC) curves of our classifier on three connectomics datasets. The classifier works best on previously unseen data of the Kasthuri volume. The dashed blue line indicates better performance on the FlyEM datasets with retraining compared to without (solid blue).}
	\label{fig:receiver-operating-characteristic}
\end{figure}

%\subsubsection{Generalization to Other Datasets}
%\noindent\textbf{Generalization to Other Datasets.}
We also evaluate the generalization to other datasets.
Our neural network does not take as input the images or affinities to allow us to transfer network weights from one dataset to the next.
Table~\TODO{X} shows the result when training on the FlyEM data and inferring on the Kasthuri data and vice versa.
The classification results are reasonable despite the fact that the datasets vary in resolution and even animal type.

@@

\begin{table}[h]
	\centering
	\small
	\setlength{\tabcolsep}{0.3em}
	\begin{tabular}{c c c c | c c c} \hline
		& \multicolumn{3}{c}{\textbf{Multicut}} & \multicolumn{3}{c}{\textbf{Lifted Multicut}} \\ \hline
		\textbf{Dataset} & $\Delta$ \textbf{Precision} & $\Delta$ \textbf{Recall} & $\Delta$ \textbf{Accuracy} & $\Delta$ \textbf{Precision} & $\Delta$ \textbf{Recall} & $\Delta$ \textbf{Accuracy} \\ \hline
		Kasthuri  &  &  & &  &  & \\
		FlyEM &  & & &  &  &  \\
		Fib-25&  & & & & & \\ \hline
	\end{tabular}
	\caption{Precision, recall, and accuracy changes between CNN only and CNN paired with graph-optimized reconstructions for the training and three test datasets. The combined method results in better precision and accuracy.}
	\label{table:multicut}
\end{table}

Our post-processing method to agglomerate segments while enforcing the acyclic nature of our graph additionally improves precision by $\TODO{XX}\%$ on average while only costing $\TODO{XX}\%$ in recall. 
Note that our post-processing strategy by design cannot increase the recall.

@@

>>>>

\subsection{Connectomics Benchmark}

In Fig.~\ref{fig:variation-of-information}, we show the VI results of the pixel-based reconstructions of the Kasthuri and FlyEM data (Sec.~\ref{sec:dataset}) for varying thresholds of agglomeration (green). 
We use one of these segmentations (green circle) as our input dataset with an agglomeration threshold of 0.3 for all datasets. 
The results from our method are shown in red for varying the $\beta$ parameter. 
We show comparisons to an oracle (blue) that correctly partitions the graph from our method based on ground truth.

Our algorithm improves the accuracy of the reconstruction for every dataset, reducing the VI split score on average by \TODO{XX}\% on the three testing datasets. 
Scores closer to the origin are better for this metric, and in every instance our results are below the green curve.
We see significant improvements on the Kasthuri datasets (VI split reduction of \TODO{XX}\%) and more modest improvements on the FlyEM datasets (reduction of \TODO{XX}\% and \TODO{XX}\%).

@@

\subsection{Empirical Ablation Studies}

\noindent\textbf{Graph Generation.}
Table \ref{table:skeletonization} shows the results of pruning the skeleton graph using the algorithm discussed in Sec.~\ref{sec:skeletonization}. 
This edge pruning is essential for the graph partitioning algorithm, which has a non-linear computational complexity dependence on the number of edges. 
The baseline algorithm considers all adjacent regions for merging. 
Our method removes a significant portion of these candidates while maintaining a large number of the true merge locations (e.g., 556 compared to 610). 
Our pruning heuristic removes at least $3.5\times$ the number of edges on all datasets, achieving a maximum removal rate of $4.8\times$.
However, there are some adjacent over-segmented labels which are not considered.

\begin{table}
	\centering
	\small
	\setlength{\tabcolsep}{0.75em}
	\begin{tabular}{c c c c c} \hline
		\textbf{Dataset} & \textbf{Segment Adjacency} & \textbf{Skeleton Pruning} & \textbf{Missed Splits} & \textbf{Gained Edges} \\ \hline
		Kasthuri & 610 / 23506 & 556 / 4921 & 214 & 160 \\
		FlyEM Vol. 1 & 304 / 15949 & 212 / 4578 & 105 & 13 \\
		FlyEM Vol. 2 &  298 / 17614 & 197 / 4366 & 120 & 19 \\ \hline
	\end{tabular}
	\caption{The results of our graph pruning approach compared to the baseline graph with all adjacent regions. We show the number of true merge locations (e.g., 610) compared to total number of edges in the graph (e.g., 23506) for each case. The number of missed splits corresponds to the number of split errors that our method misses compared to an adjacency matrix.}
	\label{table:skeletonization}
\end{table}

We generate edges in our graph by using information from the skeletons. 
In particular, we do not enforce the constraint that edges in our graph correspond to adjacent segments.
Although neurons are continuous, the EM images often have noisy spots which cause an interruption in the input segmentation.
We still want to reconstruct these neurons despite the fact that the initial segmentation is non-continuous. 
The second and fourth examples in Fig.~\ref{fig:qualitative-results} show correctly reconstructed neurons where two of the segments are non-adjacent. 
On average in our datasets we correctly merge \TODO{XX} of the non-adjacent segments.

There are some pairs of segments which we do not consider for merging because of our reliance on the skeletons.
Fig.~\ref{fig:skeleton-results} shows two such cases. 
The endpoints of both segments are circled.
In this example the small segment is carved from the larger segment in a location where there are no skeleton endpoints. 
There are on average 146 such examples in our datasets.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\linewidth]{./figures/receiver-operating-characteristic.png}		\includegraphics[width=0.45\linewidth]{./figures/precision-and-recall.png}
	\caption{The receiver operating characteristic (ROC) curves of our classifier on three connectomics datasets. The classifier works best on previously unseen data of the Kasthuri volume.}
	\label{fig:receiver-operating-characteristic}
\end{figure}

%\subsubsection{Generalization to Other Datasets}
%\noindent\textbf{Generalization to Other Datasets.}
Our neural network does not take as input the images or affinities so we can transfer network weights from one dataset to the next.
The classification results are reasonable on the FlyEM datasets despite the fact that the data varies in resolution and even animal type from the training set.

@@

\begin{table}[h]
	\centering
	\small
	\setlength{\tabcolsep}{0.3em}
	\begin{tabular}{c c c c | c c c} \hline
		& \multicolumn{3}{c}{\textbf{Multicut}} & \multicolumn{3}{c}{\textbf{Lifted Multicut}} \\ \hline
		\textbf{Dataset} & $\Delta$ \textbf{Precision} & $\Delta$ \textbf{Recall} & $\Delta$ \textbf{Accuracy} & $\Delta$ \textbf{Precision} & $\Delta$ \textbf{Recall} & $\Delta$ \textbf{Accuracy} \\ \hline
		Kasthuri  & 7.56\% & -3.83\% & 0.89\% & -5.67\% & -2.15\% & -1.03\% \\
		FlyEM Vol. 1 & 4.50\% & -4.24\% & 0.26\% & -1.52\% & -2.97\% & -0.16\%  \\
		FlyEM Vol. 2 & 5.54\% & -1.38\% & 0.26\% & -8.65\% & -2.75\% & -0.54\%\\ \hline
	\end{tabular}
	\caption{Precision, recall, and accuracy changes between CNN only and CNN paired with graph-optimized reconstructions for the training and three test datasets. The combined method results in better precision and accuracy.}
	\label{table:multicut}
\end{table}

Our post-processing method to agglomerate segments while enforcing the acyclic nature of our graph additionally improves precision by $6.6\%$ on average while only costing $-1.16\%$ in recall. 
Note that our post-processing strategy by design cannot increase the recall.

@@

Every step in our framework is fast enough on large connectomics datasets. 
After downsampling, skeletonization takes less than 1 second per segment on average. 
On average, each segment was skeletonized in 0.56 seconds on the training half of the Kasthuri data set. 
There are 4451 such segments resulting in a total running time of 2,492 seconds on an 800 megavoxel volume. 
From here, it took just over 31 seconds to generate the edges for our graph by considering nearby skeleton endpoints.
On our hardware, inference on the neural network had a throughput of 30 examples per second. 
Lastly, on this dataset the multicut algorithm ran for \TODO{XX} seconds while the lifted multicut variant spent \TODO{XX} seconds.
